### 2024-09-26 MetadataS1/MetadataS1.tex
begin table of variable format

%%%%%%% !!!! ADD DRYAD adress !!!! %%%%%%%%
Data are publicly available at \url{https://datadryad.org/}.\\[1.5em]
Raw data and codes used to extract the presented data set are publicly available at \url{https://github.com/chaireBioNorth/BYLOT_species_abundance_dataset}.
\endgroup
\newpage

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.75\linewidth]{figures/food_web.pdf} 
\caption{Synthetic vertebrate food web of the south plain of Bylot Island (figure adapted from \citet{gauthier2011}.}
\label{figure:food_web}
\end{figure}

@@

    \section*{B. Specific subproject description}
     	\subsection*{1. Site description}
 			\subsubsection*{a. Site type}
  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem in the eastern Canadian High-Arctic.An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}. The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:species_name_strategy}}; \citet{moisan2023, gauthier2024a}). The study area experiences significant temporal fluctuations in the population of small mammals (lemmings), which in turn impact the occurrence and abundance of their avian and mammalian predators such as snowy owls, rough-legged hawks, long-tailed jaegers, ermines and Arctic foxes \citep{legagneux2012}. We exclude occasional visitors, namely: i) species lacking confirmed breeding occurrences on the study site, ii) species observed solely within a single year, and iii) species primarily breeding and foraging in nearby marine or coastal habitats \citep{moisan2023}. The case of the red fox (\textit{Vulpes vulpes}) was ambiguous. While the presence of a breeding pair has been confirmed in the study area \citep{lai2022}, the extent of population establishment remains unclear and sightings are rare. Therefore, we decided to exclude this species.

@@

\begin{figure}
\centering
  \includegraphics[width=0.8\textwidth, angle=0]{figures/zones_study_area.pdf}
   \caption{Map of the different zones (colored polygons) of the 463 km\textsuperscript{2} study area located on the southern plain of Bylot Island, Nunavut Canada. The outline of the snow goose colony is presented with white dash; we used the outline of the colony in 2015 since it represents an average colony area (67 km\textsuperscript{2}).}
  \label{figure:zones}
\end{figure}

@@

		\subsubsection*{b. Avian nest monitoring}
Avian nest monitoring was conducted annually except in 2020 and 2021 due to logistical constraints imposed by the COVID-19 pandemic. Systematic nest monitoring refers here to a systematic sampling approach aimed at documenting all nests within a specified area. Monitoring is considered opportunistic when there is a chance that some nests might not have been detected within a specific area. Nest densities derived from nest sampling could be underestimated due to early nest failure (i.e., failure that happened before our sampling period).

@@

@@

   \section*{B. Accessibility}
    		\subsection*{1. Storage location and medium}
    		 Data are publicly available at \url{https://datadryad.org/}.\\[1.5em]
Raw data and codes used to extract the presented data set are publicly available at \url{https://github.com/chaireBioNorth/BYLOT_species_abundance_dataset}.

@@

        	\subsection*{4. Proprietary restrictions}
        		\subsubsection*{a. Release date} None
        		\subsubsection*{b. Citation}
        		 Please cite this data paper when using the data.
        		\subsubsection*{c. Disclaimer} None

@@

 \section*{\textcolor{Blue}{Class IV. Data structural descriptors}}
    \section*{A. Data set file}
    			\subsection*{1. Identity} 
    			a. \texttt{BYLOT-species\_taxonomy.csv}\\
     		b. \texttt{BYLOT-species\_abundance.csv}\\
     		c. \texttt{BYLOT-mean\_species\_abundance.csv}\\
     		d. \texttt{BYLOT-mean\_species\_body\_mass.csv}\\
    			e. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\

        		\subsection*{2. Size} 
        		a. 35 records, not including header row (2.3 kB)\\
        		b. 273 records, not including header row  (17.0 kB)\\
        		c. 35 records, not including header row (2.1 kB)\\
        		d. 51 records, not including header row  (1.8 kB)\\
        		e. 18 records, not including header row  (1.1 kB)

@@

        			\subsubsection*{c. \texttt{BYLOT-mean\_species\_abundance.csv}}
        			\texttt{species}; \texttt{method}; \texttt{mean\_abundance}; \texttt{sd\_abundance}; \texttt{sample\_size\_abundance}

        			\subsubsection*{d. \texttt{BYLOT-mean\_species\_body\_mass.csv}}
        			\texttt{species}; \texttt{source}; \texttt{body\_mass\_g}

        			\subsubsection*{e. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}}
        			\texttt{species}; \texttt{zone}; \texttt{mean\_nest\_density\_km2}; \texttt{sd\_nest\_density\_km2}; \texttt{sample\_size\_nest\_density\_km2}

@@

        		\subsection*{7. Authentication procedures}
        		Sums of the numeric columns: \\
        		b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2291027\\
        		c. \texttt{BYLOT-mean\_species\_abundance.csv}: mean\_abundance= 107875; sd\_abundance= 116687; sample\_size\_abundance= 207\\
        		d. \texttt{BYLOT-mean\_species\_body\_mass.csv}: body\_mass\_g= 46223\\
        		e. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}: mean\_nest\_density\_km2= 19.991; sd\_nest\_density\_km2= 10.539; sample\_size\_nest\_density\_km2= 185

>>>>

%%%%%%% !!!! ADD DRYAD and Zenodo adress !!!! %%%%%%%%
Data are publicly available at \url{https://datadryad.org/}.\\[1.5em]
Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/}.
\endgroup
\newpage

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.75\linewidth]{figures/food_web.pdf} 
\caption{Synthetic vertebrate food web of the south plain of Bylot Island.}
\label{figure:food_web}
\end{figure}

@@

    \section*{B. Specific subproject description}
     	\subsection*{1. Site description}
 			\subsubsection*{a. Site type}
  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem in the eastern Canadian High-Arctic.An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}. The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:species_name_strategy}}; \citet{moisan2023, gauthier2024a}). The study area experiences significant temporal fluctuations in the population of small mammals (lemmings), which in turn impact the occurrence and abundance of their avian and mammalian predators such as snowy owls, rough-legged hawks, long-tailed jaegers, ermines and Arctic foxes \citep{legagneux2012, duchesne2021}. We exclude occasional visitors, namely: i) species lacking confirmed breeding occurrences on the study site, ii) species observed solely within a single year, and iii) species primarily breeding and foraging in nearby marine or coastal habitats \citep{moisan2023}. The case of the red fox (\textit{Vulpes vulpes}) was ambiguous. While the presence of breeding pairs has been confirmed in the study area \citep{lai2022}, the extent of population establishment remains unclear and sightings are rare. Therefore, we decided to exclude this species.

@@

\begin{figure}
\centering
  \includegraphics[width=0.8\textwidth, angle=0]{figures/zones_study_area.pdf}
   \caption{Map of the different zones (colored polygons) of the 463 km\textsuperscript{2} study area located on the southern plain of Bylot Island, Nunavut Canada. The outline of the snow goose colony area is presented with white dash; we used the outline of the colony in 2017 since it represents an average colony area (74 km\textsuperscript{2}).}
  \label{figure:zones}
\end{figure}

@@

		\subsubsection*{b. Avian nest monitoring}
Avian nest monitoring was not conducted in 2020 and 2021 due to logistical constraints imposed by the COVID-19 pandemic. Systematic nest monitoring refers here to a systematic sampling approach aimed at documenting all nests within a specified area. Monitoring is considered opportunistic when there is a chance that some nests might not have been detected within a specific area. Nest densities derived from nest sampling could be underestimated due to early nest failure (i.e., failure that happened before our sampling period).

@@

        		\newgeometry{top=0.5cm, bottom=1.5cm, left=0.5cm, right=0.5cm}
\begin{landscape}
\input{tables/table_summary_methods.tex}
%label-> table:summary_methods
\end{landscape}
\restoregeometry
\newpage

@@

   \section*{B. Accessibility}
    		\subsection*{1. Storage location and medium}
Data are publicly available at \url{https://datadryad.org/}.\\[1.5em]
Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/}.

@@

        	\subsection*{4. Proprietary restrictions}
        		\subsubsection*{a. Release date} None
        		\subsubsection*{b. Citation}
        		 Please cite this document when using the data.
        		\subsubsection*{c. Disclaimer} None

@@

 \section*{\textcolor{Blue}{Class IV. Data structural descriptors}}
    \section*{A. Data set file}
    			\subsection*{1. Identity} 
    			a. \texttt{BYLOT-species\_taxonomy.csv}\\
     		b. \texttt{BYLOT-species\_abundance.csv}\\
     		c. \texttt{BYLOT-species\_body\_mass.csv}\\
    			d. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\

        		\subsection*{2. Size} 
        		a. 35 records, not including header row (2.3 kB)\\
        		b. 273 records, not including header row  (17.0 kB)\\
        		c. 51 records, not including header row  (1.8 kB)\\
        		d. 18 records, not including header row  (1.1 kB)

@@

        			\subsubsection*{c. \texttt{BYLOT-species\_body\_mass.csv}}
        			\texttt{species}; \texttt{source}; \texttt{body\_mass\_g}

        			\subsubsection*{d. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}}
        			\texttt{species}; \texttt{zone}; \texttt{mean\_nest\_density\_km2}; \texttt{sd\_nest\_density\_km2}; \texttt{sample\_size\_nest\_density\_km2}

@@

        		\subsection*{7. Authentication procedures}
        		Sums of the numeric columns: \\
        		b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2291027\\
        		c. \texttt{BYLOT-species\_body\_mass.csv}: body\_mass\_g= 46223\\
        		d. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}: mean\_nest\_density\_km2= 19.991; sd\_nest\_density\_km2= 10.539; sample\_size\_nest\_density\_km2= 185

### 2024-09-26 MetadataS1/tables/table_summary_variables.tex
begin table of variable format

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Sep 24 06:28:59 2024
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{longtable}{|p{0.08\textwidth}|p{0.27\textwidth}|p{0.40\textwidth}|p{0.08\textwidth}|p{0.06\textwidth}|}
\caption{Summary of variable definitions, units, and data types for each data set file.
} \\ 
  \hline
{\textbf{Data file}} & {\textbf{Variable identity}} & {\textbf{Variable definition}} & {\textbf{Units}} & {\textbf{Type}} \\ 
  \hline
a. & functional\_group & Functional group for each species. The classification of species into functional groups is based on Moisan et al. (2023). &  & string \\ 
   \hline
a. & scientific\_name & Bird scientific names were obtained from the IOC World Bird List 14.2 (Gill et al., 2024). Mammal scientific names were obtained from the Mammal species of the world: a taxonomic and geographic reference (Wilson, 2005). However, we considered here Mustela erminea richardsonii as Mustela richardsonii due to recent genetic analysis (Colella et al., 2021). &  & string \\ 
   \hline
a. & species & Common species english name. &  & string \\ 
   \hline
a. & annual\_cycle\_strategy & Annual cycle strategy of each species (i.e., resident, partially migratory or migratory). The classification of species annual cycle strategy is based on Gauthier et al., (2011)  and Moisan et al. (2023). &  & string \\ 
   \hline
b. & species & Common species english name. &  & string \\ 
   \hline
b. & year & Year corresponding to the estimate of annual abundance. & years & integer \\ 
   \hline
b. & status & Reproductive status of the individuals (i.e., undetermined or breeding). &  & string \\ 
   \hline
b. & method & Brief overview of the method used to estimate the species abundance. &  & string \\ 
   \hline
b. & abundance & Estimate of the annual number of individuals found within the 389 km2 study area located on the southern part of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. The estimates only consider adults, with the exception of lemmings, for which no distinction has been made between juveniles and adults. & individuals & integer \\ 
   \hline
c. & species & Common species english name. &  & string \\ 
   \hline
c. & method & Brief overview of the method used to estimate the species abundance. &  & string \\ 
   \hline
c. & mean\_abundance & Estimate of the mean annual number of individuals found within the 389 km2 study area located on the southern part of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. The estimates only consider adults, with the exception of lemmings, for which no distinction has been made between juveniles and adults. & individuals & integer \\ 
   \hline
c. & sd\_abundance & Standard deviation of the annual species abundance. & individuals & numeric \\ 
   \hline
c. & sample\_size\_abundance & Number of years consider in the calculation of mean annual abundance. & years & integer \\ 
   \hline
d. & species & Common species english name. &  & string \\ 
   \hline
d. & source & Source from which estimate were derived (i.e., measurements from the study area or  extracted from Wilman et al. 2014). &  & string \\ 
   \hline
d. & body\_mass\_g & Mean individual body mass. & grams & numeric \\ 
   \hline
e. & species & Common species english name. &  & string \\ 
   \hline
e. & zone & Sampled zone of the study area (see figure 2 and 3). &  & string \\ 
   \hline
e. & mean\_nest\_density\_km2 & Estimate of the mean annual nest density measured within the corresponding zone
of the study area. & nests per square kilometer & numeric \\ 
   \hline
e. & sd\_nest\_density\_km2 & Standard deviation of the annual nest density measured within the corresponding
zone of the study area. & nests per square kilometer & numeric \\ 
   \hline
e. & sample\_size\_nest\_density\_km2 & Number of years consider in the calculation of the nest density. & years & integer \\ 
   \hline
\hline
\label{table:summary_variables}
\end{longtable}
\endgroup

>>>>

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Wed Sep 25 16:53:01 2024
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{longtable}{|p{0.08\textwidth}|p{0.27\textwidth}|p{0.40\textwidth}|p{0.08\textwidth}|p{0.06\textwidth}|}
\caption{Summary of variable definitions, units, and data types for each data set file.
} \\ 
  \hline
{\textbf{Data file}} & {\textbf{Variable identity}} & {\textbf{Variable definition}} & {\textbf{Units}} & {\textbf{Type}} \\ 
  \hline
a. & functional\_group & Functional group for each species. The classification of species into functional groups is based on Moisan et al. (2023). &  & string \\ 
   \hline
a. & scientific\_name & Bird scientific names were obtained from the IOC World Bird List 14.2 (Gill et al., 2024). Mammal scientific names were obtained from the Mammal species of the world: a taxonomic and geographic reference (Wilson, 2005). However, we considered here Mustela erminea richardsonii as Mustela richardsonii due to recent genetic analysis (Colella et al., 2021). &  & string \\ 
   \hline
a. & species & Common species english name. &  & string \\ 
   \hline
a. & annual\_cycle\_strategy & Annual cycle strategy of each species (i.e., resident, partially migratory or migratory). The classification of species annual cycle strategy is based on Gauthier et al., (2011)  and Moisan et al. (2023). &  & string \\ 
   \hline
b. & species & Common species english name. &  & string \\ 
   \hline
b. & year & Year corresponding to the estimate of annual abundance. & years & integer \\ 
   \hline
b. & status & Reproductive status of the individuals (i.e., undetermined or breeding). &  & string \\ 
   \hline
b. & method & Brief overview of the method used to estimate the species abundance. &  & string \\ 
   \hline
b. & abundance & Estimate of the annual number of individuals found within the 389 km2 study area located on the southern part of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. The estimates only consider adults, with the exception of lemmings, for which no distinction has been made between juveniles and adults. & individuals & integer \\ 
   \hline
c. & species & Common species english name. &  & string \\ 
   \hline
c. & source & Source from which estimate were derived (i.e., measurements from the study area or  extracted from Wilman et al. 2014). &  & string \\ 
   \hline
c. & body\_mass\_g & Mean individual body mass. & grams & numeric \\ 
   \hline
d. & species & Common species english name. &  & string \\ 
   \hline
d. & zone & Sampled zone of the study area (see figure 2 and 3). &  & string \\ 
   \hline
d. & mean\_nest\_density\_km2 & Estimate of the mean annual nest density measured within the corresponding zone
of the study area. & nests per square kilometer & numeric \\ 
   \hline
d. & sd\_nest\_density\_km2 & Standard deviation of the annual nest density measured within the corresponding
zone of the study area. & nests per square kilometer & numeric \\ 
   \hline
d. & sample\_size\_nest\_density\_km2 & Number of years consider in the calculation of the nest density. & years & integer \\ 
   \hline
\hline
\label{table:summary_variables}
\end{longtable}
\endgroup

### 2024-09-26 manuscript/tables/table_summary_methods.tex
begin table of variable format

% Tue Sep 10 15:05:42 2024

@@

{\textbf{Species}} & {\textbf{Estimate}} & {\textbf{Method}} & {\textbf{Method quality}} & {\textbf{Justification}} & {\textbf{Min}} & {\textbf{Max}} & {\textbf{Mean}} & {\textbf{sd}} & {\textbf{n}} & {\textbf{Expert based confidence interval}} \\ 

>>>>

% Wed Sep 25 09:14:30 2024

@@

{\textbf{Species}} & {\textbf{Estimate}} & {\textbf{Method}} & {\textbf{Method quality}} & {\textbf{Justification}} & {\textbf{Lowest abundance}} & {\textbf{Highest abundance}} & {\textbf{Mean abundance}} & {\textbf{sd}} & {\textbf{n}} & {\textbf{Expert based confidence interval}} \\ 
